---
title: "RERconverge"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

PREPARING INPUT FILES (run on terminal)

Alignments from single-copy orthologues (outputs from Orthofinder):
  1. make a list of IDs based on the file names of the Single_Copy_Orthologue_Sequences folder (ls > singlecopyIDs.txt)
  2. create a folder for the alignments of interest (mkdir singlecopy_alignments)
  3. Subset in the new folder only the alignments of interest based on IDs. Run within folder with files to be copied (alignments).
        while IFS= read -r file; do cp "$file" singlecopy_alignments; done < singlecopyIDs.txt
        
```{r}
library(devtools)
library(RERconverge)
library(ape)
```

Make trees with equal topologies using phangorn. The input files are the alignments folder and the species tree (both from Orthofinder)
Before running rename species name: 
  1. In the tree file - removing _CDS_aas from each species name
  2. To change sequences IDs from all alignments to match master tree: "sed -i 's/_CDS.*//g' *" (run within folder)
```{r}
#estimatePhangornTreeAll(alndir ="NT_alignments", pattern = ".aln",
#                                          treefile = "SpeciesTree_rooted_mesquite.txt", 
#                                      output.file = "input_rerconverge.txt", submodel = "GTR",
#                                      type = "DNA", format = "fasta", k = 4)
```

```{r}
#minSpecs - minimum number of spp to calculate master tree
ort_trees <- readTrees("input_rerconverge.txt", reestimateBranches = F, minSpecs = 27) 
#estimating relative evolutionary rates (RER) - min.sp minimum number of spp to compute RER
mamRERw = getAllResiduals(ort_trees, transform = "sqrt", weighted = T, scale = T, min.sp = 16)
```

```{r}
#Foreground species
noDevDiap <- c("Frieseomelitta_varia","Tetragonula_carbonaria","Bombus_campestris","Megalopta_genalis","Halictus_rubicundus","Lasioglossum_morio","Sphecodes_monilicornis","Lasioglossum_leucozonium","Augochlora_pura","Augochlorella_aurata","Apis_mellifera","Bombus_vancouverensis","Ceratina_calcarata","Euglossa_dilemma","Melipona_quadrifasciata","Andrena_dorsata","Agapostemon_virescens")

#sister_clades = (list("clade1"=c("Bombus_vancouverensis","Bombus_campestris","Melipona_quadrifasciata","Frieseomelitta_varia","Tetragonula_carbonaria","Apis_mellifera"), "clade2" =c("Lasioglossum_leucozonium","Lasioglossum_morio","Halictus_rubicundus","Sphecodes_monilicornis","Agapostemon_virescens","Augochlora_pura","Augochlorella_aurata","Megalopta_genalis")))

#with common ancestor
#noDevDiapb2a <- foreground2TreeClades(noDevDiap, sister_clades, ort_trees)

#no ancestor
noDevDiapb2a <- foreground2Tree(noDevDiap, ort_trees, clade="terminal", weighted = "F", transition = "unidirectional")

noDevDiapplot <- plotTreeHighlightBranches(noDevDiapb2a, hlspecies = which(noDevDiapb2a$edge.length==1), hlcols="blue", main="Loss of developmental diapause")
```

```{r}
#including common ancestor
#phenVnoDevDiap <- tree2PathsClades(noDevDiapb2a,ort_trees)
#cornoDevDiap <- correlateWithBinaryPhenotype(mamRERw, phenVnoDevDiap, weighted = "auto")
#head(cornoDevDiap[order(cornoDevDiap$P),], n = 50)

#no common ancestor
phenVnoDevDiap <- tree2Paths(noDevDiapb2a,ort_trees)
cornoDevDiap <- correlateWithBinaryPhenotype(mamRERw, phenVnoDevDiap, weighted = "auto")
head(cornoDevDiap[order(cornoDevDiap$P),], n = 50)

reordered <- cornoDevDiap[order(cornoDevDiap$p.adj),]
#write.csv(reordered, file = "RER_ouput_one_per_genus.csv")

hist(cornoDevDiap$P, breaks=15)

```

```{r}
library(data.table)
terminal <- read.csv("RER_permulation_NO_ancestor.csv")
terminal
plotRers(mamRERw,"OG0005162_final_align_NT",phenv=phenVnoDevDiap)
```

Permulations
```{r}
#perform binary CC permulation
#sisters_list: a list object containing information on clades in the foreground set (pairs of sister species), specifically the pair(s) of "sister species" that branch out from the same ancestor

root_sp = "Macropis_europaea"
masterTree = ort_trees$masterTree

#with common ancestor
#permCC = getPermsBinary(1000, noDevDiap, sister_clades, root_sp, mamRERw, ort_trees, masterTree, permmode="cc")

#no common ancestor
permCC = getPermsBinary(1000, noDevDiap, sisters_list = NULL, root_sp, mamRERw, ort_trees, masterTree, permmode="cc")

#calculate permulation p-values
corpermpvals = permpvalcor(cornoDevDiap,permCC)
corpermpvals

histAlt <- hist(cornoDevDiap$P, breaks = 50)
histNull <- hist(permpvalCC, breaks = 50)

png(file="permulationSSM_sisters.png")
plot(histNull, col = "blue", xlab = "pvalues", ylab = "count")
plot(histAlt, col = "red", add = TRUE)
legend("right", inset=.02,
   c("Null(Permulations)","Alternative"), fill=c("blue","red"), horiz=TRUE, cex=0.8)
dev.off()

#add permulation to real results
reordered$permpval = corpermpvals$permpval[match(rownames(reordered), rownames(corpermpvals))]
reordered$permpbaladj <- p.adjust(reordered$permpval, method ="BH")

write.csv(reordered, file = "RER_permulation_NO_ancestor.csv")
```

TopGO
```{r}
library(topGO)

geneID2GO <- readMappings(file="TopGo_input_rn2.tsv", sep = "\t", IDsep = ",")
geneUniverse <- names(geneID2GO)
geneUniverse

cDE_cont_genes <- read.table("Rho_negative_ID.txt",header = FALSE)
cDE_cont_genes <- as.character(cDE_cont_genes$V1)

geneList <- factor(as.integer(geneUniverse%in%cDE_cont_genes))
names(geneList) <- geneUniverse

myGOdata <- new("topGOdata",description="Rho_negative", ontology="CC",allGenes=geneList, annot = annFUN.gene2GO, gene2GO = geneID2GO)

resultFisher <- runTest(myGOdata, algorithm = "weight01", statistic="fisher")
resultFisher

allRes <- GenTable(myGOdata, raw.p.value = resultFisher, classicFisher = resultFisher, orderBy = "classicFisher", ranksOf = "classicFisher", topNodes = 20)
allRes

# Kolmogorov-Smirnov testing
resultKS <- runTest(myGOdata, algorithm = "weight01", statistic = "ks")
resultKS

allRes <- GenTable(myGOdata, KS = resultKS, orderBy = "KS", ranksOf = "KS", topNodes = 20)
allRes

write.table(allRes, file = "Rho_negative_KS_CC.txt", sep = "\t", quote = FALSE, row.names = F)

#Get all the genes in your significant GO TERMS
#From: https://datacatz.wordpress.com/2018/01/19/gene-set-enrichment-analysis-with-topgo-part-1/

myterms = allRes$GO.ID 
mygenes = genesInTerm(myGOdata, myterms)
mygenes

var=c()
for (i in 1:length(myterms))
{
myterm=myterms[i]
mygenesforterm= mygenes[myterm][[1]]
mygenesforterm=paste(mygenesforterm, collapse=',')
var[i]=paste("GOTerm",myterm,"genes-",mygenesforterm)
}

write.table(var,"genesSignificanTerms.txt",sep="\t",quote=F)

#test based on gene rank

geneID2GO <- readMappings(file="TopGo_input_rn2.tsv", sep = "\t", IDsep = ",")

infile <- read.delim("RER_padj.csv", sep = ",")
infile <- na.omit(infile)
tail(infile) 

geneList <- (infile$Score)
names(geneList) <- infile$ID

topDiffGenes <- function(allScore) {
  return(allScore <= 0.055)
}
x <- topDiffGenes(geneList)
sum(x)

myGOdata <- new("topGOdata",description="Rank_test", ontology="MF",allGenes=geneList, geneSel = topDiffGenes, annot = annFUN.gene2GO, gene2GO = geneID2GO)

```
